---
title: "Attrition UU"
author: "Samuel"
date: "July 2023"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

##Introduction
introdu



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)  #The `echo = FALSE` parameter was added to the code chunk to PREVENT  printing of the R code that generated the plot.

knitr::opts_chunk$set(error = TRUE) #The `error = TRUE` parameter was added to the code chunk to PREVENT the R code to stop on error when running the code chunks.

```
# load required libraries / additional files
```{r Library,message=FALSE, echo=FALSE }

if(!require(openxlsx)){
  install.packages("openxlsx", repos = "https://www.stats.bris.ac.uk/R/")
  library(openxlsx)}


if(!require(readxl)){
  install.packages("readxl", repos = "https://www.stats.bris.ac.uk/R/")
  library(readxl)}

if(!require(corrplot)){
  install.packages("corrplot", repos = "https://www.stats.bris.ac.uk/R/")
  library(corrplot)}

if(!require(dplyr)){
  install.packages("dplyr", repos = "https://www.stats.bris.ac.uk/R/")
  library(dplyr)}


if(!require(ggplot2)){
  install.packages("ggplot2", repos = "https://www.stats.bris.ac.uk/R/")
  library(ggplot2)}

if(!require(knitr)){
  install.packages("knitr",repos = "https://www.stats.bris.ac.uk/R/")
  library(knitr)}

if(!require(class)){
  install.packages("class",repos = "https://www.stats.bris.ac.uk/R/")
  library(class)}


if(!require(lme4)){
  install.packages("lme4", repos = "https://www.stats.bris.ac.uk/R/")
  library(lme4)}

if(!require(MASS)){
  install.packages("MASS", repos = "https://www.stats.bris.ac.uk/R/")
  library(MASS)}

if(!require(tidyverse)){
  install.packages("tidyverse", repos = "https://www.stats.bris.ac.uk/R/")
  library(tidyverse)}



if(!require(assertr)){
  install.packages("assertr", repos = "https://www.stats.bris.ac.uk/R/")
  library(assertr)}

if(!require(ggpubr)){
  install.packages("ggpubr", repos = "https://www.stats.bris.ac.uk/R/")
  library(ggpubr)}

if(!require(lubridate)){
  install.packages("lubridate", repos = "https://www.stats.bris.ac.uk/R/")
  library(lubridate)}

if(!require(flextable)){
  install.packages("flextable", repos = "https://www.stats.bris.ac.uk/R/")
  library(flextable)}

if(!require(patchwork)){
    install.packages("patchwork", repos = "https://www.stats.bris.ac.uk/R/")
    library(patchwork)}


if(!require(randomForest)){
    install.packages("randomForest", repos = "https://www.stats.bris.ac.uk/R/")
    library(randomForest)}


if(!require(caret)){
    install.packages("caret", repos = "https://www.stats.bris.ac.uk/R/")
    library(caret)}


if(!require(pROC)){
    install.packages("pROC", repos = "https://www.stats.bris.ac.uk/R/")
    library(pROC)}

```



```{R}
#Set Working Directory

#setwd("Y:/15. Alteryx/Workflows/A.sam/R/Atrition")

```


```{r da}
# load dataset

#ML<- read_excel("Y:/15. Alteryx/Workflows/A.sam/R/Master Leavers.xlsx",sheet = "Leavers")

#str(ML)

#View(ML)

```


```{r}

#ML$Month<- as.Date(ML$Month, "%m/%Y")

#ML$Month <- as_datetime(ML$Month)

#ML$Month <- ymd(ML$Month)

#ML$Month <- strftime(ML$Month,format="%Y,%b")

```


```{r}

#L<- ML %>% separate(Month, c("Y","M","D"),"-")

#ML<-cbind(L$Y,L$M,ML)

#colnames(ML)[1]<- "Year"

#colnames(ML)[2]<- "Month_Year"

#View(ML)

```


***My Original Testing data***

```{r data, echo=FALSE}
# load dataset

library(readxl)

test<- read_excel("Y:/15. Alteryx/Workflows/A.sam/R/Atrition/test.xlsx")

#testdata= subset(test, select = -c(`Pers.no.`,`Name`))

testdata <- mutate_if(test, is.character, as.factor)

View(testdata)

sum(is.na(testdata))

#testdata<-na.omit(testdata)

dim(testdata)
```


***START***

```{r data, echo=FALSE}
# load dataset

library(readxl)

AT<- read_excel("Y:/15. Alteryx/Workflows/A.sam/R/Atrition/Master Leavers_.xlsx")

AT= subset(AT, select = -c(`Month`,`Leaving`))

View(AT)
dim(AT)
```


```{r}
head(AT,20)
```

**Data Check**

There are sum(is.na(AT))

```{r}
#Check if there exist missing values

sum(is.na(AT))

```

**EXPLORATORY DATA ANALYSIS** 

```{r EDA}
#total number of rows and columns 

dim(AT)

#str(AT)

```

**Groupbys'**

```{r}

by_Gender<-AT %>% group_by(Gender)


Attrition_by_day<-by_Gender %>% summarise(total_Att=sum(`Atrition`))

flextable(Attrition_by_day)

```


**Convert all columns to the proper datatype required for this analysis**
 
```{r datatype}

AT$`Month of Year` <- as.factor(AT$`Month of Year`)

#DF <- mutate_at(AT, vars(char1, char2), as.integer)

DF <- mutate_if(AT, is.character, as.factor)



str(DF)
```


**Factors Recoding**

```{r datatype}

DF<- DF%>%
  mutate(Ethnicity = recode(`Ethnic origin`,
                             "Any other Asian background" = 1,
                             "Any other Black/African/Caribbean/Black British" = 2,
                            "Any other mixed/multiple ethnic background" = 3,
                            "Arab" = 4,
                            "Asian/Asian British - Bangledeshi" = 5,
                            "Asian/Asian British - Chinese" = 6,
                            "Asian/Asian British - Indian" = 7,
                            "Asian/Asian British - Pakistani" = 8,
                            "Black/Black British - African" = 9,
                            "Black/Black British - Caribbean" = 10,
                            "Mixed - White and Asian" = 11,
                            "Mixed - White and Black African" = 12,
                            "Mixed - White and Black Caribbean" = 13,
                            "Nil Return (Not in use)" = 14,
                            "Prefer not to say" = 15,
                            "White & Other (Not in use)" = 16,
                            "White - Gypsy or Irish Traveller" = 17,
                            "White - Irish" = 18,
                            "White - Scottish  (Not in use)"= 19,
                            "White - Welsh  (Not in use)" = 20,
                           "White - any other white background" = 21,
                           "White-English/Welsh/Scottish/NorthernIrish/British" =22))




DF<- DF%>%
  mutate(Eth_Group = recode(`Ethnic Group`,
                             "Asian" = 1,
                             "Black" = 2,
                            "Mixed race" = 3,
                            "White" = 4,
                            "Not disclosed" = 5))




DF<- DF%>%
  mutate(Mar_Status = recode(`Marital Status`,
                             "Div." = 1,
                             "Marr." = 2,
                            "NM" = 3,
                            "Sep." = 4,
                            "Single" = 5,
                            "Wid." = 6,
                            "Unknwn" = 7))




DF<- DF%>%
  mutate(Cont_Type = recode(`Contract type`,
                             "Apprentice" = 1,
                             "CB - 4D" = 2,
                            "CB - Cap Prog & AMR" = 3,
                            "CB - Connections" = 4,
                            "CB - Metering" = 5,
                            "CB - Operations" = 6,
                            "CB - Ops Customer" = 7,
                            "CB - Support Service" = 8,
                            "CB - UUES" = 9,
                            "CB - UUOL" = 10,
                            "Collectively Bargain" = 11,
                            "Graduate" = 12,
                            "Local Government" = 13,
                            "Personal Contract" = 14,
                            "Trainee" = 15))


DF<- DF%>%
  mutate(Emp_Grp = recode(`Employee Group`,
                             "Permanent Employee" = 1,
                             "Temporary Employee" = 2))




DF<- DF%>%
  mutate(FuT_PaT = recode(`FT_PT`,
                             "Full Time" = 1,
                             "Part Time" = 2))





DF<- DF%>%
  mutate(Reason_For_Act_Starter = recode(`Reason For Action Starter`,
                             "External hire" = 1,
                             "Agency to Employee" = 2,
                            "TUPE Transfer" = 3))



view(DF)

```



***CORRELATION ACROSS FACTORS***

```{r COR}
DF_Corr<-DF[,c("Atrition","Leaving Month","Gen key","Age","LOS",
               "Wk.hrs.","Ethnicity","Emp_Grp","Cont_Type","Reason_For_Act_Starter", "Eth_Group","Mar_Status","FuT_PaT")]


DF_Corr<- mutate_at(DF_Corr, vars("Leaving Month"), as.integer)


cat("There is:", sum(is.na(DF_Corr)), "NA and The NA is located in row:", which(is.na(DF_Corr$Cont_Type)), "\n")

DF_Corrr<-na.omit(DF_Corr)

```



```{r Correlation}

#str(DF_Corr)

corr<-round(cor(DF_Corrr), digits = 2)

CrossTab<- c( "Atrition","Leaving.Month","Gen.key","Age","LOS",
              "Wk.hrs.","Ethnicity","Emp_Grp","Cont_Type","Reason_For_Act_Starter","Eth_Group","Mar_Status","FuT_PaT")

Cor_Table<-data.frame(CrossTab,corr)

flextable(Cor_Table)
```


```{r CorrePlot}

corrplot(corr, method = "shade", addCoef.col="black", type="upper", number.cex=0.75)

```



 **Logistic Regression MOdel (DF_Reg)**

```{r Reg_Dat}

DF_Reg<-DF[,c("Atrition","Month of Year","Age","Gender","LOS","Reason for Action","Ethnic origin", "PS group","Contract type","Employee Group","Cost Center","Position","OU L1","Age Group","Reason For Action Starter","FT_PT","Ethnic Group","Marital Status", "Ann Salary Group")]

sum(is.na(DF_Reg))

DF_Reg<-na.omit(DF_Reg)

view(DF_Reg)
```


**filter the data to remove retirement under reasons for action **

```{r Reg_Dat}

#DF_Reg <- DF_Reg %>% filter(!grepl("Retirement", `Reason for Action`))

testdata <- testdata %>% filter(!grepl("Local Government", `Contract type`))



```


**Split the data into training and testing sets**

```{r Split Data # Split the data into training and testing sets (80% training and 20% testing)}

set.seed(469820)  

train_index <- createDataPartition(DF_Reg$Atrition, p = 0.8, list = FALSE)

train_DF_Reg <- DF_Reg[train_index, ]

test_DF_Reg<- DF_Reg[-train_index, ]

```
 

**linear regression**

```{r Regression}

Lm_Reason<-lm(Atrition ~ `Reason for Action`,data =train_DF_Reg)

summary(Lm_Reason)
```


```{r Regression}

Lm_Reason<-lm(Atrition ~ `Reason For Action Starter`,data =train_DF_Reg)

summary(Lm_Reason)
```



```{r Regression}

Lm_Reason<-lm(Atrition ~ `Age Group`,data =DF_Reg)

summary(Lm_Reason)


```



```{r Regression}

Lm_Month<-lm(Atrition ~ `Month of Year`,data =train_DF_Reg)

summary(Lm_Month)
```


```{r Regression}

Lm_Los<-lm(Atrition ~ `LOS`,data =train_DF_Reg)

summary(Lm_Los)
```


```{r Regression}

Lm_CType<-lm(Atrition ~ `Contract type`,data =train_DF_Reg)

summary(Lm_CType)


```


```{r Regression}

Lm_Age<-lm(Atrition ~ `Age`,data =train_DF_Reg)

summary(Lm_Age)
```


**Model prediction**

```{r M Prediction}

P_Lm_Age<-predict(Lm_Age,newdata = test_DF_Reg)

```



```{r ROC}

roc_curve <-roc(test_DF_Reg$Atrition, as.numeric(P_Lm_Age))

plot(roc_curve, main = "ROC Curve Age", col = "blue", lwd = 4,lagacy.axes = TRUE)

text(0.5, 0.5, paste("AUC =", round(auc(roc_curve), 2)))

```



**Confusion Matrix**

```{r Confusion Matrix}

PP_Lm_Age=ifelse(test=P_Lm_Age>0.5,yes = 1,no=0)

PP_Lm_Age<- as.factor(PP_Lm_Age)

test_DF_Reg$Atrition <- as.factor(test_DF_Reg$Atrition)

Conf_matrix <- confusionMatrix(PP_Lm_Age, test_DF_Reg$Atrition)

Conf_matrix
```

**Calculate accuracy**

```{r Confusion Matrix}
accuracy <- sum(PP_Lm_Age == test_DF_Reg$Atrition) / nrow(test_DF_Reg)

cat("Accuracy:", accuracy, "\n")

```



```{r Regression}

BEN<-lm(Atrition ~ `Gender`, data = DF_Reg)

summary(BEN)
```



**Multiple linear regression**

```{r Regression}

MLM<-lm(Atrition ~ `Month of Year`+ `Age`+`Gender`+`LOS`+
           `OU L1`+`Reason for Action`+`PS group`+`Contract type`+
           `Employee Group`+`Cost Center`+`Position`+`Age Group`+
           `Reason For Action Starter`+`Ethnic Group`+`Marital Status`+`FT_PT`+
           `Ann Salary Group`, data = DF_Reg)

summary(MLM)
```


**Factor Seletion**

```{r}
stepAIC(MLM, direction = "both")
```



**Model prediction and Prediction Accuracy Check**
```{r Prediction}

P_MLM<-predict(MLM,newdata = testdata)

```
```{r ROC}

roc_curv <-roc(testdata$Atrition, as.numeric(P_MLM))

plot(roc_curv, main = "ROC Curv", col = "Green", lwd = 4,lagacy.axes = TRUE)

text(0.5, 0.5, paste("AUC =", round(auc(roc_curv), 2)))

```
**Confusion Matrix**
```{r Confusion Matrix}

P_MLM=ifelse(test=P_MLM>0.5,yes = 1,no=0)

P_MLM<- as.factor(P_MLM)

test_DF_Reg$Atrition <- as.factor(test_DF_Reg$Atrition)

C_matrix <- confusionMatrix(P_MLM, test_DF_Reg$Atrition)

C_matrix
```




*** FINAL TEST***


**Multiple linear regression: Selected Factors**

```{r Regression}

MLm<-lm(formula = Atrition ~ Age + Gender + LOS + `OU L1` + 
    `Contract type` + `Age Group` + `Reason For Action Starter` + 
    `Marital Status` + FT_PT + `Ann Salary Group`, data = DF_Reg)

#summary(MLm)

```


**Model prediction and Prediction Accuracy Check**

```{r Prediction}

P_MLm<-predict(MLm,newdata = testdata)

```


```{r ROC}

roc_curv <-roc(testdata$Atrition, as.numeric(P_MLm))

plot(roc_curv, main = "ROC Curv", col = "Green", lwd = 4,lagacy.axes = TRUE)

text(0.5, 0.5, paste("AUC =", round(auc(roc_curv), 2)))

```


**Confusion Matrix**

```{r Confusion Matrix}

P_MLm=ifelse(test=P_MLm>0.5,yes = 1,no=0)

P_MLm<- as.factor(P_MLm)

testdata$Atrition <- as.factor(testdata$Atrition)

C_matrix <- confusionMatrix(P_MLm, testdata$Atrition)

C_matrix
```




***PREDICTION**


```{r Regression}

##General prediction


Atrition<-predict(MLm,newdata = testdata)

M_dataframe <- as.data.frame(Atrition)

prediction <- cbind(testdata, M_dataframe)



#Individual Prediction


#MLm<-glm(formula = Atrition ~ Age + Gender + LOS + `OU L1` + 
    #`Contract type` + `Age Group` + `Reason For Action Starter` + 
   # `Marital Status` + FT_PT + `Ann Salary Group`, data = DF_Reg)


#*********  Age

Age<-lm(Atrition ~ `Age Group`,data =DF_Reg)

Age_m<-predict(Age,newdata = testdata)

Age_m=ifelse(test=Age_m>0.6,yes = 1,no=0)

Age_m <- as.data.frame(Age_m)

prediction<- cbind(prediction, Age_m)



#********* Gender
  
Gender<-lm(Atrition ~ `Gender`,data =DF_Reg)

Gender_m<-predict(Gender,newdata = testdata)

Gender_m=ifelse(test=Gender_m>0.6,yes = 1,no=0)

Gender_m <- as.data.frame(Gender_m)

prediction<- cbind(prediction, Gender_m)



#******* LOS

LOS<-lm(Atrition ~ `LOS`,data =DF_Reg)


LOS_m<-predict(LOS,newdata = testdata)

LOS_m=ifelse(test=LOS_m>0.6,yes = 1,no=0)

LOS_m <- as.data.frame(LOS_m)

prediction<- cbind(prediction, LOS_m)



#**OU L1**

Area<-lm(Atrition ~ `OU L1`,data =DF_Reg)

Area_m<-predict(Area, newdata = testdata)

Area_m=ifelse(test=Area_m>0.6,yes = 1,no=0)

Area_m <- as.data.frame(Area_m)

prediction<- cbind(prediction, Area_m)


#****** Ethnic_Group

Ethnic_Group <- lm(Atrition ~ `Ethnic Group`,data =DF_Reg)

EthnicGroup_m<-predict(Ethnic_Group,newdata = testdata)

EthnicGroup_m=ifelse(test=EthnicGroup_m>0.6,yes = 1,no=0)

EthnicGroup_m <- as.data.frame(EthnicGroup_m)

prediction<- cbind(prediction, EthnicGroup_m)




#******** Ann_Salary_Group

Ann_Salary_Group <- lm(Atrition ~ `Ann Salary Group`,data =DF_Reg)


AnnSalaryGroup_m<-predict(Ann_Salary_Group,newdata = testdata)

AnnSalaryGroup_m = ifelse(test= AnnSalaryGroup_m > 0.6,yes = 1,no=0)

AnnSalaryGroup_m <- as.data.frame(AnnSalaryGroup_m)

prediction<- cbind(prediction, AnnSalaryGroup_m)




#********* Marital_Status

Marital_Status<-lm(Atrition ~ `Marital Status`,data =DF_Reg)


MaritalStatus_m<-predict(Marital_Status,newdata = testdata)

MaritalStatus_m=ifelse(test=MaritalStatus_m > 0.6,yes = 1,no=0)

MaritalStatus_m <- as.data.frame(MaritalStatus_m)

prediction<- cbind(prediction, MaritalStatus_m)



#******Employee Group**

FT_PT<-lm(Atrition ~ `FT_PT`,data =DF_Reg)

FT_PT_m<-predict(FT_PT, newdata = testdata)

FT_PT_m=ifelse(test=FT_PT_m>0.6,yes = 1,no=0)

FT_PT_m <- as.data.frame(FT_PT_m)

prediction<- cbind(prediction, FT_PT_m)



#**PS group**
PSgroup<-lm(Atrition ~ `PS group`,data =DF_Reg)

PSgroup_m<-predict(PSgroup,newdata = testdata)

PSgroup_m=ifelse(test=PSgroup_m>0.6,yes = 1,no=0)

PSgroup_m <- as.data.frame(PSgroup_m)

prediction<- cbind(prediction, PSgroup_m)



#**Reason For Action Starter*

ActionType<-lm(Atrition ~ `Reason For Action Starter`,data =DF_Reg)

ActionType_m<-predict(ActionType, newdata = testdata)

ActionType_m=ifelse(test=ActionType_m>0.6,yes = 1,no=0)

ActionType_m <- as.data.frame(ActionType_m)

prediction<- cbind(prediction, ActionType_m)


#**Contract type**

Contype<-lm(Atrition ~ `Contract type`,data =DF_Reg)

Contype_m<-predict(Contype, newdata = testdata)

Contype_m=ifelse(test=Contype_m>0.6,yes = 1,no=0)

Contype_m <- as.data.frame(Contype_m)

prediction<- cbind(prediction, Contype_m)


#round up predicted attrition values to the nearest multiple of 2

prediction<- prediction %>%mutate(Atrition = abs(ceiling(abs(Atrition) * 100) / 100))

#view

View(prediction)



write.xlsx(prediction, "Y:/15. Alteryx/Workflows/A.sam/R/Atrition/prediction_October.xlsx")

```


`

if (nrow(M_dataframe) < nrow(testdata)) {
    extra_rows <- nrow(testdata) - nrow(M_dataframe)
    M_dataframe <- rbind(M_dataframe, matrix(NA, nrow = extra_rows, ncol = ncol(M_dataframe)))
}



